Factors affecting health sector expenditure in OECD countries

### 2025-02-22 code/eda-code/eda.qmd
#3

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(patchwork)
```

@@

```{r}
library(dplyr)
library(knitr)
library(kableExtra)

@@

# Create a summary table, excluding specific rows
country_summary_table <- data_2010 %>%
  filter(!CountryName %in% c("Data from database: World Development Indicators", "OECD members"
                              "Last Updated: 01/28/2025")) %>%
  select(
    CountryName,
    GDP_current_US,
    Population_total,
    Life_expectancy_at_birth_total_years,
    Physicians_per_1000_people,
    Hospital_beds_per_1000_people,
    Government_Effectiveness_Estimate,
    Control_of_Corruption_Estimate,
    Rule_of_Law_Estimate,
    Access_to_electricity__of_population,
    Unemployment_total__of_total_labor_force_modeled_ILO_estimate,
    Poverty_headcount_ratio_at_national_poverty_lines__of_population
  )

# Create and display an HTML table in R
country_summary_table %>%
  kable("html", escape = FALSE, digits = 2) %>%
  kable_styling(full_width = F, position = "left") %>%
  column_spec(1, bold = TRUE) %>%  # Making CountryName column bold
  add_header_above(c(" " = 1, "National Development Indicators" = 11)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

>>>>

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(patchwork)
library(kableExtra)
library(htmltools)
library(magrittr)
library(webshot2)
```

@@

@@

# Create a summary table, excluding specific rows
country_summary_table <- data_2010 %>%
  filter(!CountryName %in% c("Data from database: World Development Indicators", "OECD members",        "Last Updated: 01/28/2025")) %>%
  select(
    CountryName,
    GDP_current_US,
    Population_total,
    Life_expectancy_at_birth_total_years,
    Physicians_per_1000_people,
    Hospital_beds_per_1000_people,
    Government_Effectiveness_Estimate,
    Control_of_Corruption_Estimate,
    Rule_of_Law_Estimate,
    Access_to_electricity__of_population,
    Unemployment_total__of_total_labor_force_modeled_ILO_estimate,
    Poverty_headcount_ratio_at_national_poverty_lines__of_population
  )

# Generate the HTML table
table_html <- country_summary_table %>%
  kable("html", escape = FALSE, digits = 2) %>%
  kable_styling(full_width = FALSE, position = "left") %>%
  column_spec(1, bold = TRUE) %>%
  add_header_above(c(" " = 1, "National Development Indicators" = 11)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Save the table object as an RDS file
rds_file <- here::here("results", "tables", "health_exp_2010.rds")
saveRDS(table_html, file = rds_file)

### 2025-02-22 products/manuscript/Manuscript.qmd
#3

---
title: "Factors affecting health sector expenditure in OECD countries"
author: "Prasanga Paudel"
format:
  html:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
---

@@

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(here)
library(readxl)
library(writexl)
```

@@

* Prasanga Paudel$^{1,2,*}$ (ORCID: 0000-0000-1234-5678) 
* Second Author$^{3, *}$
* Third Autor$^{3, \dagger}$
* Last Author$^{1, 2, \land}$

@@

1. College of Agricultural and Environmental Science, University of Georgia, Athens, GA, USA.
2. Another Affiliation, Athens, GA, USA.
3. Yet another Affiliation.

@@

# Data import and cleaning
```{r}
file_path <- here("data","raw-data", "wdi-oecd.xlsx")
data <- read_excel(file_path, sheet="Data")
```

## A look into the original for of dataset

_This is the snapshot of the data as of now._
```{r}
head(data)
```

# EDA based on clean dataset

## Descriptive Analysis

@@

@@

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@@

@fig-health-expenditure-plot how "Health Expenditure as a Percentage of GDP Over Time (2010–2020)" is distributed over the years for top 5 OECD countries.

@@

@fig-health-exp-aus how "Health Expenditure as a Percentage of GDP changes over time for Australia.

@@

@fig-pop-growth-aus checks if Health Expenditure as a Percentage of GDP over the Years shows any reation with annual population growth for Australia.

@@

@fig-he_per_capita checks if Health Expenditure Per Capita (USD) over the Years shows any reation with annual population growth for Australia.

```{r}
#| label: fig-he_per_capita
#| fig-cap: "Health Expenditure Per Capita (USD, Country: Australia"
#| echo: FALSE
knitr::include_graphics(here("results","figures","health_exp_per_capita_aus.png"))
```

>>>>

---
title: "Factors affecting health sector expenditure in OECD countries"
author: "Prasanga Paudel"
format:
  html:
    toc: false
    embed-resources: true
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
---

@@

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(here)
library(readxl)
library(writexl)
library(magrittr)
library(patchwork)
library(kableExtra)
```

@@

* Prasanga Paudel$^{1,2,*}$ (ORCID: 0000-0000-1234-5678)

@@

1. College of Agricultural and Environmental Science, University of Georgia, Athens, GA, USA.

@@

We wll be following a panel data time series regression model. The details will be discussed as we progress.

# Importing the original dataset

```{r}
file_path <- here("data","raw-data", "wdi-oecd.xlsx")
data <- read_excel(file_path, sheet="Data")
```

## A look into the original format of the dataset

_This was the state of the data when first downloaded_
```{r}
head(data)
```

# Importing the cleaned dataset

```{r}
#Path to data. 
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

## A look into the cleaned dataset

_This is the present state of the dataset_

```{r}
head(mydata)
```

@@

In exploratory data analysis, we will just look at data from Australia as a reference for relation between variables in most of the cases. We are currently thinking of using "Current Health Expenditure as a % of GDP" as our dependent variable. But, we will see the it's distribution and connections with other variables before making decision.

@@

_Here, we will now use a combination of tables and figues to have a look into the data._

As this is a time series data, it is quite difficult to visualize all the variables across the years for all the countries. So, we will try to keep it simple and see simple relations between variables within a year or trend of a particular variable for some countries across the years.

@tbl-summary shows the values under interesting variables for all OECD countries for the year 2010.

```{r}
#| label: tbl-summary
#| tbl-cap: "Summary of variables"
#| echo: FALSE
table_html <- readRDS(here::here("results", "tables", "health_exp_2010.rds"))
table_html

```

@@

@fig-health-expenditure-plot shows how "Health Expenditure as a Percentage of GDP Over Time (2010–2020)" is distributed over the years for top 5 OECD countries.

@@

@fig-health-exp-aus shows the distribution of "Health Expenditure as a Percentage of GDP changes over time for Australia.

@@

@fig-pop-growth-aus shows the distribution  of Health Expenditure as a Percentage of GDP over the Years Comparing with shows any reation with annual population growth for Australia.This figure shows that it has very less in common with @fig-health-exp-aus as long as distribution and correlation is concerned.

@@

@fig-he_per_capita shows the distribution of Health Expenditure Per Capita (USD) over the Years. It shows that it is some what related to @fig-pop-growth-aus in it's distribution. So we can say that there is indeed a presence of some pattern in the variables of interest.

```{r}
#| label: fig-he_per_capita
#| fig-cap: "Health Expenditure Per Capita (USD, Country: Australia"
#| echo: FALSE
knitr::include_graphics(here("results","figures","health_exp_per_capita_aus.png"))
```
 I think it will be better for us to use the variable "Health Expenditure Per Capita $USD" as a dependent variable in our analysis as it show more variation and connections with other variables.
